---
title: 'Chapter 5: Control Flow'
author: "Paulo Magalang"
date: "2022-09-16"
output: 
  html_document: 
    keep_md: yes
---

# Quiz

1. What is the difference between `if` and `ifelse()`?

2. In the following code, what will the value of `y` be if `x` is `TRUE`? What if `x`
is `FALSE`? What if `x` is `NA`?

```{r}
y <- if (x) 3
```

3. What does `switch("x", x = , y = 2, z = 3)` return?

# 5.2 Choices

* `if` will return values 

```{r}
x1 <- if (TRUE) 1 else 2
x2 <- if (FALSE) 1 else 2

c(x1, x2)
```

* `if` returns `NULL` if the condition is `FALSE`

## 5.2.1 Invalid inputs

```{r, eval = FALSE}
# strange inputs to if() will throw errors
if ("x") 1
if (logical()) 1
if (NA) 1
```

```{r}
# logical vectors with length > 1 will throw a warning instead
if (c(T, F)) 1
```

```{r, eval = FALSE}
# can turn this behavior into an error by setting an environment var
Sys.setenv("_R_CHECK_LENGTH_1_CONDITION_" = "true")
if (c(TRUE, FALSE)) 1
```

## 5.2.2 Vectorised if

* `ifelse(test, yes, no)`

* `dplyr::case_when()` useful instead of nested `ifelse()` statements

```{r}
x <- 1:10
dplyr::case_when(
  x %% 35 == 0 ~ "fizz buzz",
  x %% 5 == 0 ~ "fizz",
  x %% 7 == 0 ~ "buzz",
  is.na(x) ~ "???",
  TRUE ~ as.character(x)
)
```

## 5.2.3 `switch()` statement

```{r}
# instead of:
x_option <- function(x) {
  if (x == "a") {
    "option 1"
  } else if (x == "b") {
    "option 2" 
  } else if (x == "c") {
    "option 3"
  } else {
    stop("Invalid `x` value")
  }
}

# can use a switch statement to make the code more compact
x_option <- function(x) {
  switch(x,
    a = "option 1",
    b = "option 2",
    c = "option 3",
    stop("Invalid `x` value") # last component should throw an error since 
                              # unmatched inputs will return NULL
  )
}
```

```{r}
# similar to C's switch statements, outputs will "fall through" to next value
legs <- function(x) {
  switch(x,
    cow = , # 4
    horse = , # 4
    dog = 4,
    human = , # 2
    chicken = 2,
    plant = 0,
    stop("Unknown input")
  )
}
legs("cow")
legs("dog")
```

* recommend to use `switch()` with character inputs

## 5.2.4 Exercises

1. What type of vector does each of the following calls to `ifelse()` return?

```{r}
ifelse(TRUE, 1, "no") # double
ifelse(FALSE, 1, "no") # character
ifelse(NA, 1, "no") # logical, NA is logical by default
```

2. Why does the following code work?

```{r}
x <- 1:10 # length of x != 0
if (length(x)) "not empty" else "empty"

x <- numeric() # length of x = 0
if (length(x)) "not empty" else "empty"

# booleans can be interpreted as 1 (for TRUE) and 0 (for FALSE)
# I'm assuming if() expects any integer input and will output TRUE for any nonzero value

# will if() be happy with double input?

if(3.2) 1 else 0 # double inputs are rounded down to the nearest whole integer
if(0.1) 1 else 0
if(-1) 1 else 0 # negative values are nonzero
```
















