---
title: 'Chapter 17: Big Picture'
author: "Paulo Magalang"
date: "2023-06-29"
output: html_document
---

```{r}
library(rlang)
library(lobstr)
```


## 17.2 Code is data

* capture code with `rlang::expr()`; captured code are expressions

```{r}
expr(mean(x, na.rm = TRUE))
expr(10 + 100 + 1000)
```

* need to use `enexpr()` to capture code passed to a function

```{r}
capture_it <- function(x) {
  enexpr(x)
}
capture_it(a + b + c)
```

```{r}
f <- expr(f(x = 1, y = 2))

# Add a new argument
f$z <- 3
f

# Or remove an argument:
f[[2]] <- NULL
f
```

## 17.3 Code is a tree

* code is represented as an abstract syntax tree (AST)

* can use `lobstr::ast()` to inspect tree

```{r}
lobstr::ast(f(a, "b"))
```

```{r}
lobstr::ast(f1(f2(a, b), f3(1, f4(2))))
```

```{r}
lobstr::ast(1 + 2 * 3)
```

## 17.4 Code can generate code

* `call2()` constructs a function call from the function to call and its arguments

```{r}
call2("f", 1, 2, 3)
call2("+", 1, call2("*", 2, 3))
```

* `!!` (bang-bang) is the unquote operator

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

expr(!!xx / !!yy)
```

```{r}
cv <- function(var) {
  var <- enexpr(var)
  expr(sd(!!var) / mean(!!var))
}

cv(x)
cv(x + y)

cv(`)`)
```

## 17.5 Evaluation runs code

```{r}
eval(expr(x + y), env(x = 1, y = 10))
eval(expr(x + y), env(x = 2, y = 100))
```

```{r}
x <- 10
y <- 100
eval(expr(x + y))
```

## 17.6 Customizing evaluation with functions

```{r}
string_math <- function(x) {
  e <- env(
    caller_env(),
    `+` = function(x, y) paste0(x, y),
    `*` = function(x, y) strrep(x, y)
    # create environment where + and * works with strings
  )

  eval(enexpr(x), e)
}

name <- "Hadley"
string_math("Hello " + name)
string_math(("x" * 2 + "-y") * 3)
```

## 17.7 Customizing evaluation with data

```{r}
df <- data.frame(x = 1:5, y = sample(5))
eval_tidy(expr(x + y), df)
```

```{r}
with2 <- function(df, expr) {
  eval_tidy(enexpr(expr), df)
}

with2(df, x + y)
```

## 17.8 Quosures

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enexpr(expr), df)
}
```

```{r}
df <- data.frame(x = 1:3)
a <- 10
with2(df, x + a) # a in the environment is 1000
```

```{r}
with2 <- function(df, expr) {
  a <- 1000
  eval_tidy(enquo(expr), df)
}

with2(df, x + a)
```



















