---
title: 'Chapter 13: S3'
author: "Paulo Magalang"
date: "2023-03-02"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(sloop)
```

# 13.2: Basics

* S3 objects is abase type with a `class` attribute minimum

```{r}
f <- factor(c("a", "b", "c"))

typeof(f)
attributes(f)
```

```{r}
# unclass strips the class atribute
unclass(f)
```

* S3 objects behave differently from underlying base tyeps when passed to a generic;
use `sloop::ftype()` to determine if a function is a generic

```{r}
ftype(print)
ftype(str)
ftype(unclass)
```

* generics define an interface which depends on the class of the argument; example of method
dispatch

```{r}
print(f)

# stripping class reverts to integer behaviour
print(unclass(f))
```

```{r}
time <- strptime(c("2017-01-01", "2020-05-04 03:21"), "%Y-%m-%d")
str(time)

str(unclass(time))
```

```{r}
# see the process of method dispatch
s3_dispatch(print(f))
```

* methods usually have a `.` in the function name but some base R functions also
have `.`'s in the function name

```{r, eval = FALSE}
# use sloop::s3_get_method() to get source code from S3 methods
weighted.mean.Date
s3_get_method(weighted.mean.Date)
```

## 13.2.1 Exercises

1. Describe the difference between `t.test()` and `t.data.frame()`. When is each
function called?

```{r}
ftype(t.test)
ftype(t.data.frame)

s3_dispatch(print(t.test))
s3_dispatch(print(t.data.frame))
```

2. Make a list of commonly used base R functions that contain `.` in their name but
are not S3 methods.

* `data.frame`

* `read.csv`

* `install.packages`

3. What does the `as.data.frame.data.frame()` method do? Why is it confusing? How could you
avoid this confusion in your own code?

`as.data.frame.data.frame()` checks to see if the object is a data frame and converts it into
a data frame if it is not. It is confusing since the function call contains many `.`. It can
be rewritten as `as.data.frame(x)` where `x` is the object of interest.

4. Describe the difference in behavior in these two calls.

```{r}
set.seed(1014)
some_days <- as.Date("2017-01-31") + sample(10, 5)

some_days
mean(some_days) # returns a date object that gives the mean date
mean(unclass(some_days)) # stripped class behavior for it to behave like numeric values
```

5. What class of object does the following code return? What base type is it built on?
What attributes does it use?

```{r}
x <- ecdf(rpois(100, 10))
#x

ftype(x)
typeof(x)
attributes(x) # returns ecdf object, `call` attribute stores function call
```

6. What class of object does the following code return? What base type is it built on?
What attributes does it use?

```{r}
x <- table(rpois(100, 5))
#x

typeof(x)
attributes(x) # returns table obj, stores table names in `dimnames`
```

# 13.3 Classes

* can directly set class attributes to make an object an instance of a class

```{r}
# Create and assign class in one step
x <- structure(list(), class = "my_class")

# Create, then set class
x <- list()
class(x) <- "my_class"
```


```{r}
# determine S3 class with class()
class(x)

# check if an object is an instance of a class with inherits()
inherits(x, "my_class")
inherits(x, "your_class")
```

```{r}
# Create a linear model
mod <- lm(log(mpg) ~ log(disp), data = mtcars)
class(mod)
print(mod)

# Turn it into a date (?!)
class(mod) <- "Date"

# Unsurprisingly this doesn't work very well
#print(mod)
```

* constructor: `new_myclass()` that creates new objects with correct structure

* validator: `validate_myclass()` that performs checks to ensure object has correct values

* helper: `myclass()` that provides a convenient way for others to create objects of the class


## 13.3.1 Constructors

* constructors enforce consistent object structure

* 3 principles: 

1. be called `my_newclass()`

2. have one argument for the base object and one for each attribute

3. check the type of base object and the types of each attribute

```{r}
new_Date <- function(x = double()) {
  stopifnot(is.double(x))
  structure(x, class = "Date")
}

new_Date(c(-1, 0, 1))
```

```{r}
new_difftime <- function(x = double(), units = "secs") {
  stopifnot(is.double(x))
  units <- match.arg(units, c("secs", "mins", "hours", "days", "weeks"))

  structure(x,
    class = "difftime",
    units = units
  )
}

new_difftime(c(1, 10, 3600), "secs")
new_difftime(52, "weeks")
```


## 13.3.2 Validators

* constructors only checks if types are correct, better to validate elsewhere to make
initializing the object cheap

```{r}
new_factor <- function(x = integer(), levels = character()) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))

  structure(
    x,
    levels = levels,
    class = "factor"
  )
}

# malformed factors
#new_factor(1:5, "a")
#new_factor(0:1, "a")
```

```{r}
validate_factor <- function(x) {
  values <- unclass(x)
  levels <- attr(x, "levels")

  if (!all(!is.na(values) & values > 0)) {
    stop(
      "All `x` values must be non-missing and greater than zero",
      call. = FALSE
    )
  }

  if (length(levels) < max(values)) {
    stop(
      "There must be at least as many `levels` as possible values in `x`",
      call. = FALSE
    )
  }

  x
}

#validate_factor(new_factor(1:5, "a"))
#validate_factor(new_factor(0:1, "a"))
```

## 13.3.3 Helpers

* helpers always have the same name as the class

* finish by calling the constructor and validator

* create meaningful error messages and UX

```{r}
difftime <- function(x = double(), units = "secs") {
  x <- as.double(x) # coercion in helper, not constructor
  new_difftime(x, units = units)
}

difftime(1:10)
```

```{r}
factor <- function(x = character(), levels = unique(x)) {
  ind <- match(x, levels) # guess that the levels should be unique values
  validate_factor(new_factor(ind, levels))
}

factor(c("a", "a", "b"))
```

```{r}
POSIXct <- function(year = integer(), 
                    month = integer(), 
                    day = integer(), 
                    hour = 0L, 
                    minute = 0L, 
                    sec = 0, 
                    tzone = "") {
  ISOdatetime(year, month, day, hour, minute, sec, tz = tzone)
}

POSIXct(2020, 1, 1, tzone = "America/New_York")
```

## 13.3.4 Exercises

1. Write a constructor for `data.frame` objects. What base type is a data frame built on?
What attributes does it use? What are the restrictions placed on individual elements?
What about the names?

```{r}
df <- data.frame(x = c(1, 2), y = c(3, 4))
typeof(df)
attributes(df)
```


```{r}
new_data.frame <- function(x = list(), n = numeric(), row.names = NULL) {
  stopifnot(is.list(x))
  
  # need to check if all columns are the same length
  stopifnot(unique(all(lengths(x))) == 1)
  
  if(is.null(row.names)) { # if no row.names given, set row names as integers
    row.names <- 1:n
  } else {
    # need to check if row.names are characters
    stopifnot(is.character(row.names))
    
    # also if the user supplied the correct number of rows
    stopifnot(length(row.names) == n)
    
    # row names must be unique
    stopifnot(length(row.names) == length(unique(row.names)))
  }
  
  structure(x,
            class = "data.frame",
            row.names = row.names)
}
```

```{r}
x <- list(a = 1, b = 2)
new_data.frame(x, n = 1)
```

2. Enhance my `factor()` helper to have better behavior when one or more `values` is not
found in `levels`. What does `base::factor()` do in this situation?

```{r}
base::factor(c("a", "b", "c"), levels = c("a", "b"))
```

```{r}
new_factor <- function(x = integer(), levels = character()) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))

  structure(
    x,
    levels = levels,
    class = "factor"
  )
}
```

```{r}
factor <- function(x = character(), levels = unique(x)) {
  ind <- match(x, levels) 
  
  # just throw an error if the user-supplied level is not defined
  if (sum(is.na(ind)) > 0) { # if NA found in ind
    missing_levels <- setdiff(x, levels)
    stop("factors ", paste0("'", missing_levels, "'", collapse = ", "), "not defined")
  }
  
  
  validate_factor(new_factor(ind, levels))
}

factor(c("a", "a", "b"))
```

```{r}
x <- c("a", "b", "c")
levels <- c("a", "b")
match(x, levels)
```

3. Carefully read the source code of `factor()`. What does it do that my constructor does
not?

```
function (x = character(), levels, labels = levels, exclude = NA, 
    ordered = is.ordered(x), nmax = NA) 
{
    if (is.null(x)) 
        x <- character() # forces input to be character
    nx <- names(x)
    if (missing(levels)) { # automatically set levels based on x if not supplied
        y <- unique(x, nmax = nmax)
        ind <- order(y)
        levels <- unique(as.character(y)[ind])
    }
    force(ordered)
    if (!is.character(x)) 
        x <- as.character(x)
    levels <- levels[is.na(match(levels, exclude))]
    f <- match(x, levels)
    if (!is.null(nx)) 
        names(f) <- nx
    if (missing(labels)) {
        levels(f) <- as.character(levels)
    }
    else {
        nlab <- length(labels)
        if (nlab == length(levels)) {
            nlevs <- unique(xlevs <- as.character(labels))
            at <- attributes(f)
            at$levels <- nlevs
            f <- match(xlevs, nlevs)[f]
            attributes(f) <- at
        }
        else if (nlab == 1L) 
            levels(f) <- paste0(labels, seq_along(levels))
        else stop(gettextf("invalid 'labels'; length %d should be 1 or %d", 
            nlab, length(levels)), domain = NA)
    }
    class(f) <- c(if (ordered) "ordered", "factor")
    f
}
```

4. Factors have an optimal "contrasts" attribute. Read the help for `C()`, and
briefly describe the purpose of the attribute. What type should it have? Rewrite
the `new_factor()` constructor to include this attribute.

The help file for `C()` was not that informative but it did mention that it can
be a matrix or a character string.

```{r}
new_factor <- function(x = integer(), levels = character(), contrasts = NULL) {
  stopifnot(is.integer(x))
  stopifnot(is.character(levels))
  
  if(!is.null(contrasts)) { # contrasts can be either matrix or character
    stopifnot(is.matrix(contrasts) | is.character(contrasts))
  }

  structure(
    x,
    levels = levels,
    class = "factor",
    contrasts = contrasts
  )
}
```

5. Read the documentation for `utils::as.roman()`. How would you write a constructor
for this class? Does it need a validator? What might a helper do?

```{r}
new_roman <- function(x = integer()) {
  stopifnot(is.integer(x))
  structure(x, class = "roman")
}
```

```{r}
# documentation states that numbers between 1 and 3899 have unique representation as roman numerals
validate_roman <- function(x) {
  value <- unclass(x)
  
  if (value > 3899 | value < 1) {
    stop("Only numbers between 1 and 3899 have a unique representation as roman numbers")
  }
  
  x
}
```

```{r}
roman <- function(x = numeric()) {
  x <- as.integer(x)
  
  validate_roman(new_roman(x))
}
```

```{r}
x <- roman(1)
x
```









































