---
title: 'Chapter 9: Functionals'
author: "Paulo Magalang"
date: "2023-01-17"
output: 
  html_document: 
    keep_md: yes
---

# 9.1 Introduction

* functional: takes a function as an input and returns a vector as output

```{r}
set.seed(1)

randomise <- function(f) f(runif(1e3))
randomise(mean)
randomise(mean)
randomise(sum)
```

* common examples: R apply family functions, purrr `map()`

# 9.2 My first functional: `map()`

```{r}
library(purrr)
```

* `map(1:3, f)` equivalent to `list(f(1), f(2), f(3))`

```{r}
# map(vector, function)

triple <- function(x) x * 3
map(1:3, triple)
```

```{r}
# basic implementation of map()

simple_map <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```

* R base equivalent to `purrr::map()` is `lapply()`


## 9.2.1 Producing atomic vectors

```{r}
# map_chr() always returns a character vector
map_chr(mtcars, typeof)

# map_lgl() always returns a logical vector
map_lgl(mtcars, is.double)

# map_int() always returns a integer vector
n_unique <- function(x) length(unique(x))
map_int(mtcars, n_unique)

# map_dbl() always returns a double vector
map_dbl(mtcars, mean)
```

```{r}
pair <- function(x) c(x, x)
map(1:2, pair)
map(1:2, as.character)
```

* `sapply()` and `vapply()` also outputs atomic vectors

## 9.2.2 Anonymous functions and shortcuts

```{r}
# inline anonymous function
map_dbl(mtcars, function(x) length(unique(x)))
```

```{r}
# purrr shortcut
map_dbl(mtcars, ~ length(unique(.x)))
```

```{r}
as_mapper(~ length(unique(.x)))
```

```{r}
# extracting elements using map()

x <- list(
  list(-1, x = 1, y = c(2), z = "a"),
  list(-2, x = 4, y = c(5, 6), z = "b"),
  list(-3, x = 8, y = c(9, 10, 11))
)

# Select by name
map_dbl(x, "x")

# Or by position
map_dbl(x, 1)

# Or by both
map_dbl(x, list("y", 1))

# You'll get an error if a component doesn't exist:
#map_chr(x, "z")

# Unless you supply a .default value
map_chr(x, "z", .default = NA)
```

## 9.2.3 Passing arguments with `...`

```{r}
x <- list(1:5, c(1:10, NA))
map_dbl(x, ~ mean(.x, na.rm = TRUE))
```

```{r}
# directly pass na.rm to mean(), no need to specify parameters in function call
map_dbl(x, mean, na.rm = TRUE)
```

* arguments passed this way are not decomposed! (ie: passing a vector as an argument
will stay as a vector)

```{r}
plus <- function(x, y) x + y
x <- c(0, 0, 0, 0)

# passing extra arguments in map()
map_dbl(x, plus, runif(1)) # runif() evaluated once

# passing extra arguments in anonymous function
map_dbl(x, ~ plus(.x, runif(1))) # runif() evaluated for each index
```

## 9.2.4 Argument names

```{r}
# can write out full argument names passed through map()
x <- list(1:5, 5:10)
map(x, mean, trim = 0.1)
```

```{r}
simple_map <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```

```{r}
# arguments to map() are .x and .f
boostrap_summary <- function(x, f) {
  f(sample(x, replace = TRUE))
}

#simple_map(mtcars, bootstrap_summary, f = mean)
# really simple_map(x = mtcars, f = mean, bootstrap_summary), bootstrap_summary passed as trim in mean()
```

## 9.2.5 Varying another argument

* How to keep first argument of `map()` constant and vary the argument?

```{r}
trims_param <- c(0, 0.1, 0.2, 0.5)
vect <- rcauchy(1000)
```

* manually rearrange argument order using an anonymous function

```{r}
map_dbl(trims_param, ~ mean(vect, trim = .x))
```

```{r}
# more explicit
map_dbl(trims_param, function(trim) mean(vect, trim = trim))
```

* take advantage of argument matching rules

```{r}
#map_dbl(trims, mean, x = x)
# this is confusing. no reason to do this
```

## 9.2.6 Exercises

1. Use `as_mapper()` to explore how purrr generates anonymous functions for the integer, character, and list
helpers. What helper allows you to extract attributes? Read the documentation to find out.

```{r}
as_mapper(1)
```

```{r}
as_mapper("a")
```

```{r}
as_mapper(list(1, "a"))
```

If the input is a character vector, numeric vector, or list, it is converted to an
extractor function in `purrr::pluck()`.

2. `map(1:3, ~ runif(2))` is a useful pattern for generating random numbers, but
`map(1:3, runif(2))` is not. Why not? Can you explain why it returns the result
that it does?

`~ runif(2)` is executed for each value in `1:3` while `runif(2)` is executed only
one time BUT why are vector values returned?

```{r}
map(1:3, ~ runif(2))
# map(1:3, function(x) runif(2))
```

```{r}
as_mapper(~ runif(2))
```

```{r}
map(1:3, runif(2))
```

```{r}
as_mapper(runif(2))
```

```{r, eval = FALSE}
v <- 1:3
rand <- as.list(runif(2)) # rand will always be a double < 1

# from as_mapper(runif(2)) output, I would expect two values to be output since
# runif(2) will generate two random values
# also pluck() behavior
pluck(v[1], rand[[1]], .default = NULL)
pluck(v[1], rand[[2]], .default = NULL)
# input and output vectors must have same dimensions and indices < 1 are treated as 1 as opposed to regular vector behavior
# returning empty vectors

v[rand[[1]]]

# but executing pluck() verbosely from as_mapper() throws an error since v[1] and rand dims are not equal
pluck(v[1], rand, .default = NULL)
#pluck(v, rand, .default = NULL)

# also pluck_raw() doesn't exist, probably a C wrapper idk
```

3. Use the appropriate `map()` function to:

a. Compute the standard deviation of every column in a numeric data frame.

```{r}
map_dbl(mtcars, ~ sd(.x))
```

b. Compute the standard deviation of every numeric column in a mixed data frame.
(Hint: you'll need to do it in two steps.)

```{r}
library(tidyverse)
mtcars_mixed <- rownames_to_column(mtcars, "name")

#map_dbl(mtcars_mixed, ~ sd(.x)) # will introduce NAs

# determine which columns are numeric
num_col <- map_lgl(mtcars_mixed, is.numeric)

# subset numeric columns and calculate sd
map_dbl(mtcars_mixed[num_col], ~ sd(.x))
```

c. Compute the number of levels for every factor in a data frame.

```{r}
factor_col <- map_lgl(iris, is.factor)
map_int(iris[factor_col], ~ length(levels(.x)))
```

4. The following code simulates the performance of a t-test for non-normal data.
Extract the p-value from each test, then visualize.

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))
p_vals <- map_dbl(trials, "p.value")

boxplot(p_vals, col = "white", horizontal = T)
stripchart(p_vals, method = "jitter", pch = 19, add = TRUE)
```

5. The following code uses a map nested inside another map to apply a function to every
element of a nested list. Why does it fail, and what do you need to do to make it work?

```{r}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
# map(x, map, .f = triple)
# map is passed to triple() as an argument

#map(x, ~ map(triple(.x)))
map(x, ~ map(.x, triple))
```

6. Use `map()` to fit linear models to the `mtcars` dataset using the formulas stored in
this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

fit_models <- map(formulas, lm, data = mtcars)
```

7. Fit the model `mpg ~ disp` to each of the bootstrap replicates of `mtcars` in the
list below, then extract the $R^2$ of the model fit (Hint: you can compute the $R^2$
with `summary()`.)

```{r}
set.seed(1)
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))
mod <- mpg ~ disp
mod_fit <- map(bootstraps, ~ lm(mod, data = .x))

# summarize across and extract r.squared
map(map(mod_fit, summary), "r.squared")

# one liner possible but will look ugly or just pipe (section 9.3)
```

# 9.3 Purrr style

* just pipe map using `%>%` (or R base functions) to make code more readable

```{r}
by_cyl <- split(mtcars, mtcars$cyl)

by_cyl %>% 
  map(~ lm(mpg ~ wt, data = .x)) %>% 
  map(coef) %>% 
  map_dbl(2)
```

# 9.4 Map variants

* output same type as input with `modify()`

* iterate over two inputs with `map2()`

* iterate with index using `imap()`

* return nothing with `walk()`

* iterate over any number of inputs with `pmap()`

## Same type of output as input: `modify()`

* `map()` always returns a list, use `modify()` to return the same type of output
as the input

```{r}
df <- data.frame(
  x = 1:3,
  y = 6:4
)

map(df, ~ .x * 2)
modify(df, ~ .x * 2)
```

* `modify()` returns modified copy

```{r}
simple_modify <- function(x, f, ...) {
  for (i in seq_along(x)) {
    x[[i]] <- f(x[[i]], ...)
  }
  x
}
```

## Two inputs: `map2()` and friends

```{r}
xs <- map(1:8, ~ runif(10))
xs[[1]][[1]] <- NA
ws <- map(1:8, ~ rpois(10, 5) + 1)

map_dbl(xs, mean) # unweighted means

# can't pass ws as an additional argument
#map_dbl(xs, weighted.mean, w = ws)
```

```{r}
# map2() is vectorized over two arguments, .x and .y to each call to .f
map2_dbl(xs, ws, weighted.mean)
```

```{r}
# basic implementation
simple_map2 <- function(x, y, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], y[[i]], ...)
  }
  out
}

# actual map2() will recycle inputs!
```

## 9.4.3 No outputs: `walk()` and friends

* not all functions are needed for their output values but for their side effects
(`cat()`, `write.csv()`, `ggsave()`)

```{r}
welcome <- function(x) {
  cat("Welcome ", x, "!\n", sep = "")
}
names <- c("Hadley", "Jenny")

# As well as generate the welcomes, it also shows 
# the return value of cat()
map(names, welcome)
```

```{r}
# walk() ignores the return values of .f and returns .x invisibly
walk(names, welcome)
```

## 9.4.4 Iterating over values and indices

```{r}
imap_chr(iris, ~ paste0("The first value of ", .y, " is ", .x[[1]]))
```

```{r}
x <- map(1:6, ~ sample(1000, 10))
imap_chr(x, ~ paste0("The highest value of ", .y, " is ", max(.x)))
```

## 9.4.5 Any number of inputs: `pmap()` and friends

* `pmap(list(x, y, z, etc...), f, additional_args)`

* `tibble::tribble()` allows user to describe a dataframe row-by-row

```{r}
params <- tibble::tribble(
  ~ n, ~ min, ~ max,
   1L,     0,     1,
   2L,    10,   100,
   3L,   100,  1000
)

pmap(params, runif)
# runif(n, min, max)
```

* `pmap()` base equivalents: `Map()` and `mapply()`

## 9.4.6 Exercises

1. Explain the results of `modify(mtcars, 1)`.

Since `mtcars` is a dataframe, the output will also be a dataframe. The `1` refers to the
first row of data in `mtcars` and is mapped to every row of `mtcars` which means the 32 rows
of the output will reflect the first row of `mtcars`

```{r}
modify(mtcars, 1)
mtcars[1, ]
dim(mtcars)
```


2. Rewrite the following code to use `iwalk()` instead of `walk2()`. What are
the advantages and disadvantages?

```{r}
cyls <- split(mtcars, mtcars$cyl)
temp <- "./"
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
walk2(cyls, paths, write.csv)
```

```{r}
# imap(x,f) is equivalent to map2(x, names(x), f)
cyls <- split(mtcars, mtcars$cyl)
temp <- "./"
names(cyls) <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
iwalk(cyls, write.csv)
```


3. Explain how the following code transforms a data frame using functions
stored in a list.

```{r}
trans <- list(
  disp = function(x) x * 0.0163871,
  am = function(x) factor(x, labels = c("auto", "manual"))
)

nm <- names(trans) # only mutate on disp and am cols
mtcars[nm] <- map2(trans, mtcars[nm], function(f, var) f(var)) # for each function in trans, mutate on disp and am cols
                                                               # in mtcars; execute function on specific col
```

Compare and contrast the `map2()` approach to this `map()` approach:

```{r}
mtcars[nm] <- map(nm, ~ trans[[.x]](mtcars[[.x]]))
                        # analogous to function(f, var) f(var)
# trans[[.x]] either disp() or am() FUNCTION
# mtcars[[.x]] refers to actual mtcars col
```

```{r}
as_mapper(~ trans[[.x]](mtcars[[.x]]))
```

4. What does `write.csv()` return, i.e. what happens if you use it with `map2()` instead
of `walk2()`?

Returns `NULL`

```{r}
cyls <- split(mtcars, mtcars$cyl)
temp <- "./"
paths <- file.path(temp, paste0("cyl-", names(cyls), ".csv"))
map2(cyls, paths, write.csv)
```

# 9.5 Reduce family

## 9.5.1 Basics

* `reduce(1:4, f)` equivalent to `f(f(f(1, 2), 3), 4)`, useful for sequential function calls

```{r}
l <- map(1:4, ~ sample(1:10, 15, replace = T))
str(l)
```

```{r}
out <- l[[1]]
out <- intersect(out, l[[2]])
out <- intersect(out, l[[3]])
out <- intersect(out, l[[4]])
out
```

```{r}
# instead use reduce
reduce(l, intersect)
```

```{r}
reduce(l, union)
```

## 9.5.2 Accumulate

* `accumulate()` returns intermediary results also

```{r}
accumulate(l, intersect)
```

```{r}
x <- c(4, 3, 10)
reduce(x, `+`)

accumulate(x, `+`)
```

## 9.5.3 Output types

* `.init` arg in `reduce()` defines initial value to start accumualtion; also helpful
to ensure valid input/output types

```{r}
reduce(integer(), `+`, .init = 0)
#reduce("a", `+`, .init = 0)
```

## 9.5.4 Multiple inputs

* `reduce2(.x, .y, f, .init)`

* $|.y|$ must be $|.x| - 1$ if `.init` is not specified; if `.init` is supplied,
both magnitudes can be the equal

## 9.5.5 Map-reduce

* map-reduce: map combined with reduce split between computers, divide and conquer

# 9.6 Predicate functionals

* predicate returns T/F

## 9.6.1 Basics

* predicate functional applies predicate to each element of a vector

* `some(.x, .p)` returns T if ANY element matches

* `every(.x, .p)` returns T if ALL elements match

* `none(.x, .p)` returns T if NO element matches

* `detect(.x, .p)` returns the value of the first match; `detect_index(.x, .p)` returns
index of first match

* `keep(.x, .p)` keeps all matching elements; `discard(.x, .p)` drops all matching
elements

## 9.6.2 Map variants

```{r}
df <- data.frame(
  num1 = c(0, 10, 20),
  num2 = c(5, 6, 7),
  chr1 = c("a", "b", "c"),
  stringsAsFactors = FALSE
)

str(map_if(df, is.numeric, mean))
str(modify_if(df, is.numeric, mean))
str(map(keep(df, is.numeric), mean))
```

## 9.6.3 Exercises

1. Why isn't `is.na()` a predicate function? What base function is closest to being a
predicate version of `is.na()`?

`is.na()` is not a predicate function because the test is applied to the entire vector and
returns a logical vector. `anyNA()` returns one T/F value.

2. `simple_reduce()` has a problem when `x` is length 0 or 1. Describe the source of the
problem and how you might go about fixing it.

```{r}
simple_reduce <- function(x, f) {
  out <- x[[1]]
  for (i in seq(2, length(x))) {
    out <- f(out, x[[i]])
  }
  out
}
```

```{r}
simple_reduce <- function(x, f) {
  if (length(x) == 0) {
    stop("Error: `.x` is empty.")
  } 
  else if (length(x) == 1) {
    return(x)
  }
  else {
    out <- x[[1]]
    for (i in seq(2, length(x))) {
      out <- f(out, x[[i]])
    }
    return(out)
  }
}
```

```{r}
x <- c(1, 2, 3)
simple_reduce(x, `+`)
simple_reduce(1, `+`)
simple_reduce("a", `+`)
#simple_reduce(integer(), `+`)
```

3. Implement the `span()` function from Haskell: given a list `x` and a predicate
function `f`, `span(x, f)` returns the location of the longest sequential run of
elements where the predicate is true. (Hint: you might find `rle()` helpful.)

```{r}
test <- list(
  c(999, 0, 0, 0, 999, 999),
  c(0, 0, 0, 999, 999, 999, 0, 999),
  c(999, 999, 0, 0, 0, 0, 0, 0, 0)
)

test_2max <- list(
  c(999, 0, 0, 0, 999, 999),
  c(0, 0, 0, 999, 999, 0, 999),
  c(999, 999)
)

is_zero <- function(value) {
  value == 0
}



runs <- map(test, is_zero) %>% map(rle)
runs
```

```{r}
# split runs into lengths and values
length <- runs %>% map("lengths")
value <- runs %>% map("values")

length
value
```

```{r}
#map(length, ~ which(.x == max(.x))) # returns index of longest runs but cannot differentiate between T/F values

# take advantage of map2
map2(length, value, ~ which(.x == max(.x) & .y == TRUE))
# finds longest runs in each vector but not overall list

#map(length, max) %>% unlist
```


4. Implement `arg_max()`. It should take a function and a vector of inputs, and return
the elements of the input where the function returns the highest value. For example,
`arg_max(-10:5, function(x) x ^ 2)` should return -10. `arg_max(-5:5, function(x) x ^ 2)`
should return `c(-5, 5)`. Also implement the matching `arg_min()` function.

```{r}
arg_max <- function(vect, f) {
  values <- f(vect)
  
  # from values, determine index of max number
  max_index <- which(values == max(values))
  
  return(vect[max_index])
}

arg_max(-5:5, function(x) x ^ 2)
```

```{r}
arg_max <- function(vect, f) {
  values <- f(vect)
  
  max_index <- which(values == min(values))
  
  return(vect[max_index])
}

arg_max(-5:5, function(x) x ^ 2)
```

5. The function below scales a vector so it falls in the range [0, 1]. How would you apply
it to every column of a data frame? How would you apply it to every numeric column in a
data frame?

```{r}
scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

df <- data.frame(x = 1:3,
                 y = c("a", "b", "c"),
                 z = 4:6,
                 a = c(T, F, F))

map_if(df, is.numeric, scale01)
```

# 9.7 Base functionals

## 9.7.1 Matrices and arrays

* `base::apply()` to work with two-dimensional and higher vectors while `map()` works best
with 1D vectors

```{r}
a2d <- matrix(1:20, nrow = 5)
apply(a2d, 1, mean)
apply(a2d, 2, mean)
```

```{r}
a3d <- array(1:24, c(2, 3, 4))
apply(a3d, 1, mean)
apply(a3d, c(1, 2), mean)
```

* caveats of using `apply()`: no control of output type, not idempotent (output not always
same of input)


## 9.7.2 Mathematical concerns

* `integrate()` finds the area under the curve defined by `f()`

* `uniroot()` finds where `f()` hits zero

* `optimize()` finds the location of the lowest (or highest) value of `f()`

```{r}
integrate(sin, 0, pi)
str(uniroot(sin, pi * c(1 / 2, 3 / 2)))
str(optimise(sin, c(0, 2 * pi)))
str(optimise(sin, c(0, pi), maximum = TRUE))
```

## 9.7.3 Exercises

1. How does `apply()` arrange the output? Read the documentation and perform some experiments.

```{r}
a2d <- matrix(1:20, nrow = 5)
a2d
apply(a2d, 1, mean)
apply(a2d, 2, mean)
```

2. What do `eapply()` and `rapply()` do? Does purrr have equivalents?

`eapply()` applies a function over values in an environment. `rapply()` recursively
applies a function over a list.

3. Challenge: read about the fixed point algorithm. Complete the exercises using R.













