---
title: 'Chapter 7: Environments'
author: "Paulo Magalang"
date: "2022-10-18"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rlang)
```

# 7.1 Introduction

## Quiz

1. List at least three ways that an environment differs from a list?

2. What is the parent of the global environment? What is the only environment
that doesn't have a parent?

3. What is the enclosing environment of a function? Why is it important?

4. How do you determine the environment from which a function was called?

5. How are `<-` and `<<-` different?

# 7.2 Environment basics

* environments are similar to a named list except:

  - every name is unique
  
  - names are unordered
  
  - has a parent
  
  - no copy on modify
  
## 7.2.1 Basics

```{r}
# create environments

# with rlang::env
e1 <- env(
  a = FALSE,
  b = "a",
  c = 2.3,
  d = 1:3,
)

# envrionments can contain themselves
e1$d <- e1

# with base::new.env (from 1st edition of the book)
e <- new.env() # init environment
e$a <- FALSE
e$b <- "a"
e$c <- 2.3
e$d <- 1:3
```

```{r}
# memory addresses
e1 # just prints address
env_print(e1) # has more info
```

```{r}
# print names
env_names(e1)
names(e1)
ls(e1, all.names = TRUE)
```

## 7.2.2 Important environments

* current environment `current_env()`: environment where code is currently executing

* global environment `global_env()`: environment where interactive computation takes place

* access global env with `globalenv()` and current env with `environment()`

```{r}
# environment comparison: use identical()
identical(global_env(), current_env())

#global_env() == current_env()
```

## 7.2.3 Parents

* every environment has a parent environment, used to implement lexical scoping

```{r}
e2a <- env(d = 4, e = 5)
e2b <- env(e2a, a = 1, b = 2, c = 3)
```

```{r}
# find parent of an environment with env_parent()
env_parent(e2b)
env_parent(e2a)
```

* empty environment does not have a parent

```{r}
e2c <- env(empty_env(), d = 4, e = 5)
e2d <- env(e2c, a = 1, b = 2, c = 3)
```

```{r}
# can see ancestors with env_parents()
env_parents(e2b) # by default stops at global environment
env_parents(e2d) # ancestors terminate at empty environment
```

```{r}
# ancestors of global environment includes all attached packages
env_parents(e2b, last = empty_env())
```

* `parent.env()` finds parent of environment

## 7.2.4 Super assignment `<<-`

* `<-` creates a variable in the current environment, `<<-` modifies an existing variable
in a parent env

* if a variable is not found using `<<-`, one is created in the global environment

```{r}
x <- 0
f <- function() {
  x <<- 1
}
f()
x
```

## 7.2.5 Getting and setting

* get and set elements of an environment with `$` and `[[`, but cannot use numeric
indices

```{r}
e3 <- env(x = 1, y = 2)
e3$x
e3$z <- 3
e3[["z"]]
```

* if binding does not exist, will return `NULL` but can change default value in
`env_get()` or just throw an error

* `env_poke()` takes a name and value to add additional bindings

```{r}
env_poke(e3, "a", 100)
e3$a
```

* `env_bind` to add multiple bindings

```{r}
env_bind(e3, a = 10, b = 20)
env_names(e3)
```

* `env_has()` to check if binding exists in an env

* `env_unbind()` to unbind objects from env

```{r}
e3$a <- NULL
env_has(e3, "a")

env_unbind(e3, "a")
env_has(e3, "a")
```

* be careful with `get`, `assign`, `exists`, and `rm`: designed to be used interactively

## 7.2.6 Advanced bindings

* `env_bind_lazy()` creates delayed bindings - bindings that are evaluated the first
time they are accessed, useful for accessing data from R packages

```{r}
env_bind_lazy(current_env(), b = {Sys.sleep(1); 1}) # b not evaluated

system.time(print(b)) # now evaluated
system.time(print(b)) # already loaded in, no need to reevaluate
```

* `env_bind_active()` creates active bindings - bindings that are recomputed every
time they are accessed, used for R6 active fields

```{r}
env_bind_active(current_env(), z1 = function(val) runif(1))

z1
z1
```

## 7.2.7 Exercises

1. List three ways in which an environment differs from a list.

* environments not copied on modification, cannot be subset using numerical
indices, and each environment has a parent

2. Create an environment as illustrated by this picture.

```{r}
loop_env <- env()
loop_env$loop <- loop_env
```

3. Create a pair of environments as illustrated by this picture.

```{r}
deloop_env <- env()
loop_env <- env()

loop_env$loop <- deloop_env
deloop_env$deloop <- loop_env
```

4. Explain why `e[[1]]` and `e[c("a", "b")]` don't make sense when `e` is an environment.

The book mentioned that environments cannot be subset using numeric indices. This
makes sense knowing that environment names are not ordered, unlike lists. 

I'm not sure why accessing multiple environment names would not work.

5. Create a version of `env_poke()` that will only bind new names, never re-bind
old names. Some programming languages only do this, and are known as single assignment
languages.

```{r}
env_poke_SA <- function(environment, name, value) {
    # check to see if binding in the environment exists
    if(!env_has(environment, name)) { # if binding does not exist
        environment[[name]] <- value # create binding
    }
}

# test with existing e3
env_names(e3)
e3$x

env_poke_SA(e3, "x", 2)
e3$x
```

6. What does this function do? How does it differ from `<<-` and why might you
prefer it?

`rebind()` recursively checks through each parent environment to see if `name` exists.
If `name` is found in a parent environment, it is reassigned `value`. If not, it will
check the subsequent parent environment up until the empty environment. `<<-` only
stops at the global environment, not the empty environment, and will create a new
binding in the global environment if the name does not exist.

```{r}
rebind <- function(name, value, env = caller_env()) {
  if (identical(env, empty_env())) { # check to see if we went back to far, base case
    stop("Can't find `", name, "`", call. = FALSE)
  } else if (env_has(env, name)) {
    env_poke(env, name, value) # reassign value when it exists, terminate recursion
  } else {
    rebind(name, value, env_parent(env)) # recursive call, reduction step
  }
}

#rebind("a", 10)
a <- 5
rebind("a", 10)
a
```

# 7.3 Recursing over environments

* `where()` looks for a name starting at a given environment

```{r}
where <- function(name, env = caller_env()) {
  if (identical(env, empty_env())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
  } else if (env_has(env, name)) {
    # Success case
    env
  } else {
    # Recursive case
    where(name, env_parent(env))
  }
}
```

```{r}
#where("yyy")

x <- 5
where("x")

where("mean")
```

## 7.3.1 Exercises

1. Modify `where()` to return all environments that contain a binding for `name`.
Carefully think through what type of object the function will need to return.

```{r}
# kind of a cheat solution. function doesn't explicitly returns an output but
# concatenates to an existing list

env_list <- c() # initialize empty output list
where_all <- function(name, env = caller_env()) {
  if (identical(env, empty_env())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
  } else if (env_has(env, name)) {
    # Success case
    env_list <<- c(env_list, env)
    # need another recursive call after success case, and rethink termination
  } else {
    # Recursive case
    where_all(name, env_parent(env))
  }
}

# test
e4a <- env(empty_env(), a = 1, b = 2)
e4a

e4b <- env(e4a, x = 10, a = 11)
e4b

where_all("a", env = e4b)
env_list
```




























