---
title: 'Chapter 14: R6'
author: "Paulo Magalang"
date: "2023-03-31"
output: html_document
---

# 14.1: Introduction

* R6 objects are encapsulated and called as `object$method()`; objects
are also mutable

* similar to base OOP system called reference classes

```{r}
library(R6)
```

# 14.2 Classes and methods

* `R6Class(classname, public)`

* `classname` is optional but nice to have, use camel case by convention

* `public`: list of methods and fields that make up the public interface
of the object, snake case by convention; refer to methods and fields of current
instance via `self$`

```{r}
Accumulator <- R6Class("Accumulator", list(
  sum = 0,
  add = function(x = 1) {
    self$sum <- self$sum + x 
    invisible(self)
  })
)
```


```{r}
Accumulator
```

* initialize instance of object by calling `new()`

```{r}
x <- Accumulator$new()
```

```{r}
x$add(4) 
x$sum
```

## 14.2.1 Method chaining

```{r}
x$add(10)$add(10)$sum

# for readability
x$
  add(10)$
  add(10)$
  sum
```

## 14.2.2 Important methods

* important methods that should be defined for all classes: `$initialize()` and `$print()`

```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) { # overrides $new()
    stopifnot(is.character(name), length(name) == 1)
    stopifnot(is.numeric(age), length(age) == 1)
    
    self$name <- name
    self$age <- age
  }
))

#hadley <- Person$new("Hadley", age = "thirty-eight")

hadley <- Person$new("Hadley", age = 38)
```


```{r}
Person <- R6Class("Person", list(
  name = NULL,
  age = NA,
  initialize = function(name, age = NA) {
    self$name <- name
    self$age <- age
  },
  print = function(...) { # overrides default printing behavior
    cat("Person: \n")
    cat("  Name: ", self$name, "\n", sep = "")
    cat("  Age:  ", self$age, "\n", sep = "")
    invisible(self)
  }
))

hadley2 <- Person$new("Hadley")
hadley2
```

## 14.2.3 Adding methods after creation

* add elements to class with `$set()`

```{r}
Accumulator <- R6Class("Accumulator")
Accumulator$set("public", "sum", 0)
Accumulator$set("public", "add", function(x = 1) {
  self$sum <- self$sum + x 
  invisible(self)
})
```

## 14.2.4 Inheritance

```{r}
AccumulatorChatty <- R6Class("AccumulatorChatty", 
  inherit = Accumulator,
  public = list(
    add = function(x = 1) {
      cat("Adding ", x, "\n", sep = "")
      super$add(x = x)
    }
  )
)

x2 <- AccumulatorChatty$new()
x2$add(10)$add(1)$sum
```

## 14.2.5 Introspection

```{r}
class(hadley2)
names(hadley2)
```

## 14.2.6 Exercises

1. Create a bank account R6 class that stores a balance and allows you to deposit and withdraw
money. Create a subclass that throws an error if you attempt to overdraft. Create another
subclass that allows you to go into overdraft, but charges you a fee.

```{r}
BankAccount <- R6Class("BankAccount", public = list(
  balance = 0,
  
  initialize = function(balance = 0) {
    stopifnot(is.numeric(balance))
    self$balance <- balance
  },
  
  deposit = function(deposit) {
    self$balance <- self$balance + deposit
    invisible(self)
  },
  
  withdraw = function(withdraw) {
    self$balance <- self$balance - withdraw
    invisible(self)
  },
  
  print = function(...) {
    cat("Current balance: ", self$balance)
  })
)
```

```{r}
my_acct <- BankAccount$new(100)
my_acct$deposit(10)$balance
my_acct
```

```{r}
BankAccountStrict <- R6Class("BankAccountStrict",
  inherit = BankAccount,
  public = list(
    withdraw = function(withdraw) {
      if(self$balance - withdraw < 0) {
        stop("You are broke.")
      }
      super$withdraw(withdraw)
    }
  )
)
```

```{r}
broke_acct <- BankAccountStrict$new(10)
#broke_acct$withdraw(100)
```

```{r}
BankAccountOverdraft <- R6Class("BankAccountOverdraft",
  inherit = BankAccount,
  public = list(
    withdraw = function(withdraw) {
      if(self$balance - withdraw < 0) {
        message("$10 charge added for overdraft.")
      }
      super$withdraw(withdraw + 10)
    }
  )
)
```

```{r}
overdraft_acct <- BankAccountOverdraft$new(0)
overdraft_acct$withdraw(100)
overdraft_acct
```

2. Create an R6 class that represents a shuffled deck of cards. You should be able to
draw cards from the deck with `$draw(n)` and return all cards to the deck and reshuffle
with `$reshuffle()`. Use the following code to make a vector of cards.

```{r}
suit <- c("♠", "♥", "♦", "♣")
value <- c("A", 2:10, "J", "Q", "K")
cards <- paste0(rep(value, 4), suit)
length(cards)
```

```{r}
Deck <- R6Class("Deck", public = list(
  cards = NULL,
  randomizer = NULL,
  index = 1,
  initialize = function(...) {
    suit <- c("♠", "♥", "♦", "♣")
    value <- c("A", 2:10, "J", "Q", "K")
    cards <- paste0(rep(value, 4), suit)
    
    self$cards <- cards
    self$randomizer <- sample(1:52, size = 52, replace = F)
  },
  draw = function(n) {
    # check if there are enough cards
    if (self$index + n - 1 > 52) {
      stop("There are not enough cards.")
    }
    
    hand <- cards[randomizer[self$index:(self$index + n - 1)]]
    self$index <- self$index + n
    hand
  },
  reshuffle = function(...) {
    self$randomizer <- sample(1:52, size = 52, replace = F)
    self$index <- 1
    invisible(self)
  })
)
```

3. Why can't you model a bank account or a deck of cards with an S3 class?

R6 objects are mutable while S3 objects exhibit copy on modify behavior.

4. Create an R6 class that allows you to get and set the current time zone. You
can access the current time zone with `Sys.timezone()` and set with `Sys.setenv(TZ = "newtimezone")`.
When setting the time zone, make sure the new time zone is in the list provided by `OlsonNames()`

```{r}
TimeZone <- R6Class("TimeZone", public = list(
  current_timezone = NULL,
  initialize = function(tz = Sys.timezone()) {
    self$current_timezone <- tz
  },
  get_timezone = function(...) {
    cat("Current time zone:", self$current_timezone)
  },
  set_timezone = function(tz) {
    # validation
    stopifnot(is.character(tz), tz %in% OlsonNames())
    
    Sys.setenv(TZ = tz)
    self$current_timezone <- tz
    invisible(self)
  })
)
```


5. Create an R6 class that manages the current working directory. It should have
`$get()` and `$set()` methods.

```{r}
WorkingDirectory <- R6Class("WorkingDirectory", public = list(
  get = function(...) {
    getwd()
  },
  set = function(dir) {
    stopifnot(dir.exists(dir), is.character(dir))
    setwd(dir)
  })
)
```

6. Why can't you model the time zone or current working directory with an S3 class?

7. What base type are R6 objects built on top of? What attributes do they have?


# 14.3 Controlling access

* `private`: create fields and methods that are available within the class

* `active`: use accessor functions to define dynamic fields

## 14.3.1 Privacy

* `private` argument works similarly to `public`

* fields and methods are accessed using `private$` instead of `self$`

```{r}
Person <- R6Class("Person", 
  public = list(
    initialize = function(name, age = NA) {
      private$name <- name
      private$age <- age
    },
    print = function(...) {
      cat("Person: \n")
      cat("  Name: ", private$name, "\n", sep = "")
      cat("  Age:  ", private$age, "\n", sep = "")
    }
  ),
  private = list(
    age = NA,
    name = NULL
  )
)

hadley3 <- Person$new("Hadley")
hadley3
hadley3$name # would be Hadley if was not a private field
```

## 14.3.2 Active fields

* active fields define components that look like fields but are defined with functions;
implemented with active bindings

```{r}
Rando <- R6::R6Class("Rando", active = list(
  random = function(value) {
    if (missing(value)) {
      runif(1)  
    } else {
      stop("Can't set `$random`", call. = FALSE)
    }
  }
))
x <- Rando$new()
x$random
x$random
x$random
```

* active fields useful with private fields

```{r}
Person <- R6Class("Person", 
  private = list(
    .age = NA,
    .name = NULL
  ),
  active = list(
    age = function(value) {
      if (missing(value)) {
        private$.age
      } else {
        stop("`$age` is read only", call. = FALSE)
      }
    },
    name = function(value) {
      if (missing(value)) {
        private$.name
      } else {
        stopifnot(is.character(value), length(value) == 1)
        private$.name <- value
        self
      }
    }
  ),
  public = list(
    initialize = function(name, age = NA) {
      private$.name <- name
      private$.age <- age
    }
  )
)

hadley4 <- Person$new("Hadley", age = 38)
hadley4$name
#hadley4$name <- 10
hadley4$name <- "Had"
hadley4$name

#hadley4$age <- 20
```

## 14.3.3 Exercises

1. Create a bank account class that prevents you from directly setting the account balance,
but you can still withdraw from and deposit to. Throw an error if you attempt to go into
overdraft.

```{r}
BankAccountPrivate <- R6Class("BankAccountPrivate", 
  public = list(
    initialize = function(balance = 0) {
      private$balance <- balance
    },
    deposit = function(dep) {
      private$balance <- private$balance + dep
      invisible(self)
    },
    withdraw = function(draw) {
      if (private$balance - draw < 0) {
        stop("You are broke.")
      }
      private$balance <- private$balance - draw
      invisible(self)
    }
  ),
  private = list(
    balance = NULL
  )
)
```

2. Create a class with a write-only `$password` field. It should have `$check_password(password)`
method that returns `TRUE` or `FALSE`, but there should be no way to view the complete password.

```{r}
Password <- R6Class("Password",
  public = list(
    initialize = function(password) {
      if (missing(password)) {
        stop("Password missing.")
      } else {
        private$password <- password
      }
    },
    check_password = function(password) {
      return(private$password == password)
    }
  ),
  private = list(
    password = NULL
  )
)
```

3. Extend the `Rando` class with another active binding that allows you to access the previous
random value. Ensure that active binding is the only way to access the value.

```{r}
Rando <- R6::R6Class("Rando", 
  active = list(
    random = function(value) {
      if (missing(value)) {
        if (is.na(private$current_value)) {
          private$current_value <- runif(1)
          private$current_value
        } else {
          private$prev_value <- private$current_value
          private$current_value <- runif(1)
          private$current_value
        }
      } else {
        stop("Can't set `$random`", call. = FALSE)
      }
    },
    prev_random = function() {
      if (missing(value)) {
        private$prev_value
      } else {
        stop("Can't set `$prev_random`", call. = FALSE)
      }
    }
  ),
  private = list(
    prev_value = NA,
    current_value = NA
  )
)
```

4. Can subclasses access private fields/methods from their parent? Perform an experiment
to find out.

```{r}
Class <- R6::R6Class("Class",
  private = list(
    private_field = "This is the private field.",
    private_method = function() {
      "This is the private method."
    }
  )
)

OtherClass <- R6::R6Class("OtherClass",
  inherit = Class,
  public = list(
    return_superclass_priv = function() {
      priv_field <- super$private_field
      priv_method <- super$private_method()
      
      cat(priv_field)
      cat(priv_method)
    }
  )
)

OtherClass$new()$return_superclass_priv()
```

# 14.4 Reference semantics

* R6 objects are not copied when modified

```{r}
y1 <- Accumulator$new() 
y2 <- y1

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)
```

```{r}
y1 <- Accumulator$new() 
y2 <- y1$clone() # use $clone() to copy an object

y1$add(10)
c(y1 = y1$sum, y2 = y2$sum)
```

## 14.4.1 Reasoning

```{r, eval = FALSE}
x <- list(a = 1)
y <- list(b = 2)

z <- f(x, y)
```

```{r, eval = FALSE}
x <- List$new(a = 1)
y <- List$new(b = 2)

z <- f(x, y)
# write functions that return a value or modify R6 inputs, not both simultaneously
```

## 14.4.2 Finalizer

* `$finalize()` complements `$initialize()`; similar role to `on.exit()`

```{r}
TemporaryFile <- R6::R6Class("TemporaryFile", list(
  path = NULL,
  initialize = function() {
    self$path <- tempfile()
  },
  finalize = function() {
    message("Cleaning up ", self$path)
    unlink(self$path)
  }
))
```

```{r}
tf <- TemporaryFile$new()
rm(tf)
```

## 14.4.3 R6 fields

* if an R6 class is used as a default value of a field, it will be shared across
all instances of the object

```{r}
TemporaryDatabase <- R6::R6Class("TemporaryDatabase", list(
  con = NULL,
  file = TemporaryFile$new(), # same instance
  initialize = function() {
    self$con <- DBI::dbConnect(RSQLite::SQLite(), path = file$path)
  },
  finalize = function() {
    DBI::dbDisconnect(self$con)
  }
))

db_a <- TemporaryDatabase$new()
db_b <- TemporaryDatabase$new()

db_a$file$path == db_b$file$path
```

```{r}
TemporaryDatabase <- R6::R6Class("TemporaryDatabase", list(
  con = NULL,
  file = NULL,
  initialize = function() {
    self$file <- TemporaryFile$new() # new instance created
    self$con <- DBI::dbConnect(RSQLite::SQLite(), path = file$path)
  },
  finalize = function() {
    DBI::dbDisconnect(self$con)
  }
))

db_a <- TemporaryDatabase$new()
db_b <- TemporaryDatabase$new()

db_a$file$path == db_b$file$path
```

## 14.4.4 Exercises

1. Create a class that allows you to write a line to a specified file. You should
open a connection to the file in `$initialize()`, append a line using `cat()` in
`$append_line()`, and close the connection in `$finalize()`.

```{r}
AppendFile <- R6::R6Class("AppendFile", 
  public = list(
    file = NA,
    intialize = function(filename) {
      self$file <- open(filename)
    },
    appendline = function(string) {
      if (is.na(self$file)) {
        stop("No open connection to file.")
      }
      cat(string, file = self$file, append = TRUE)
    },
    finalize = function() {
      close(self$file)
    }
)
```



























