---
title: "Object Oriented Programming Intro and Base Types"
author: "Paulo Magalang"
date: "2023-03-01"
output: 
  html_document: 
    keep_md: yes
---

# Object-oriented Programming Introduction

* OOP systems: S3, R6, S4

* S3 and S4 use generic function OOP, not encapsulated OOP (aka python, java)

* functional programming more important in R since it uses generic function OOP

## OOP systems

* polymorphism: function's interface is seperate from implementation

```
// simple example of static polymorphism in java
// example from https://stackify.com/oop-concept-polymorphism/

// create a CoffeeMachine object and a method to brew coffees
// how many coffees to brew? one or many?

public class BasicCoffeeMachine {
    public Coffee brewCoffee(CoffeeSelection selection) throws CoffeeException {
    // brewCoffee() implementation to brew only one selection of coffee
        switch (selection) {
        case FILTER_COFFEE:
            return brewFilterCoffee();
        default:
            throw new CoffeeException(
                "CoffeeSelection ["+selection+"] not supported!");
        }   
    }
  
    public List brewCoffee(CoffeeSelection selection, int number) throws CoffeeException {
    // brewCoffee() implementation to brew multiple cups of coffee defined by number
        List coffees = new ArrayList(number);
        for (int i=0; i<number; i++) {
            coffees.add(brewCoffee(selection));
        }
        return coffees;
    }
}
```

* example of polymorphism in R

```{r}
diamonds <- ggplot2::diamonds

summary(diamonds$carat) # diamonds$carat is doubles
summary(diamonds$cut) # diamonds$cut is factors
```

* classes define the object (like the fields of an object), methods define what objects can do

* method dispatch defines how methods are inherited from parent classes

* encapsulated OOP: methods belong to objects or classes, `object.method(arg1, arg2)`

* functional OOP: methods belong to generic functions, `generic(object, arg2, arg3)`


## OOP in R

* S3: informal implementation of functional OOP

* S4: formal and rigorous rewrite of S3; provides more guarantees and encapsulation

* RC: encapsulated OO; special type of S4 objects that are mutable

* R6: encapsulated OOP like RC but resolves some issues described in 14.5

* R.oo: mutable S3 objects

* proto: blurs distinctions between classes and instances of classes


## sloop

```{r}
library(sloop)

otype(1:10)
otype(mtcars)
mle_obj <- stats4::mle(function(x = 1) (x - 2) ^ 2)
otype(mle_obj)
```

# 12. Base Types

## 12.2 Base versus OO objects

```{r}
# A base object:
is.object(1:10)
sloop::otype(1:10)

# An OO object
is.object(mtcars)
sloop::otype(mtcars)
```

```{r}
# OO object have "class" attribute
attr(1:10, "class")
attr(mtcars, "class")
```

```{r}
# class() safe for S3 and S4 objects, but misleading for base objects
x <- matrix(1:4, nrow = 2)
class(x)
sloop::s3_class(x)
```

## 12.3 Base types

```{r}
# every object has a base type
typeof(1:10)
typeof(mtcars)
```

* vectors: `NULL`, `logical`, `integer`, `double`, `complex`, `character`, `list`, and `raw`

```{r}
typeof(NULL)
typeof(1L)
typeof(1i)
```

* functions: `closure`, `special`, `builtin`

```{r}
typeof(mean)
typeof(`[`)
typeof(sum)    
```

* environments: `environment`

```{r}
typeof(globalenv())
```

* `S4`: S4 classes that don't inherit from base type

```{r}
mle_obj <- stats4::mle(function(x = 1) (x - 2) ^ 2)
typeof(mle_obj)
```

* language components: `symbol`, `language`, `pairlist`, `expression`

```{r}
typeof(quote(a))
typeof(quote(a + 1))
typeof(formals(mean))
```

* rarely seen in R: `externalptr`, `weakref`, `bytecode`, `promise`, `...`, `any`

### 12.3.1 Numeric type

* 3 different meanings of "numeric"

1. alias for double type

2. in S3/S4, numeric means either integer or double type

```{r}
sloop::s3_class(1)
sloop::s3_class(1L)
```

3. `is.numeric()` tests for objects that behave like numbers, ex: factors

```{r}
typeof(factor("x"))
is.numeric(factor("x"))
```




