---
title: 'Chapter 15: S4'
author: "Paulo Magalang"
date: "2023-04-20"
output: 
  html_document: 
    keep_md: yes
---

# 15.1 Introduction

* slot: a named component of S4 objects, use `@` as subsetting operator

* no one reference that has all S4 QandA; sometimes clashes with base R best
practices

```{r}
library(methods) # available for most of the time, but no harm no foul
```

# 15.2 Basics

* define S4 classes with `setClass()`

```{r}
setClass("Person", 
  slots = c(
    name = "character", 
    age = "numeric"
  )
)
```

* construct new objects with `new()`

```{r}
john <- new("Person", name = "John Smith", age = NA_real_)

is(john)
john@name
slot(john, "age")
```

* try not to use `@` when using other people's classes, look for accessor functions
(getters and setters)

```{r}
# create getter and setter
setGeneric("age", function(x) standardGeneric("age"))
setGeneric("age<-", function(x, value) standardGeneric("age<-"))

# define those methods with setMethod()
setMethod("age", "Person", function(x) x@age)
setMethod("age<-", "Person", function(x, value) {
  x@age <- value
  x
})

age(john) <- 50
age(john)
```

```{r}
sloop::otype(john)
sloop::ftype(age)
```

## 15.2.1 Exercises

1. `lubridate::period()` returns an S4 class. What slots does it have? What class
is each slot? What accessors does it provide?

```{r}
sloop::otype(lubridate::period())

str(lubridate::period())
#methods?lubridate::period()
```

2. What other ways can you find help for a method? Read `?"?"` and summarize the details.

* `methods?generic` # look for all methods

* `method?generic(args)` # docs for specific method


# 15.3 Classes

* need three args for `setClass()`: name, slots, prototype

```{r}
setClass("Person", 
  slots = c(
    name = "character", 
    age = "numeric"
  ), 
  prototype = list(
    name = NA_character_,
    age = NA_real_
  )
)

me <- new("Person", name = "Hadley")
str(me)
```

## 15.3.1 Inheritance

* `contains`: specifies a class to inherit slots and behavior from

```{r}
setClass("Employee", 
  contains = "Person", 
  slots = c(
    boss = "Person"
  ),
  prototype = list(
    boss = new("Person")
  )
)

str(new("Employee"))
```

## 15.3.2 Introspetion

* use `is()` to determine what classes an object inehrits from

```{r}
is(new("Person"))
is(new("Employee"))
is(john, "Person")
```

## 15.3.3 Redefinition

```{r}
setClass("A", slots = c(x = "numeric"))
a <- new("A", x = 10)

#setClass("A", slots = c(a_different_slot = "numeric"))
#a
```

## 15.3.4 Helper

* helper functions: have same name as class; UX centered; finish by calling
`methods::new()`

```{r}
Person <- function(name, age = NA) {
  age <- as.double(age)
  
  new("Person", name = name, age = age)
}

Person("Hadley")
```

## 15.3.5 Validator

```{r}
Person("Hadley", age = c(30, 37)) # age is vector length 2; doesnt make sense
```

```{r}
# enforce using validator
setValidity("Person", function(object) {
  if (length(object@name) != length(object@age)) {
    "@name and @age must be same length"
  } else {
    TRUE
  }
})
```

```{r}
# Person("Hadley", age = c(30, 37))
```

## 15.3.6 Exercises

1. Extend the person class with fields to match `utils::person()`. Think about what
slots you need, what class each shot should have, and what you'll need to check
in your validity method.

```{r}
setClass("Person2", 
  slots = c(
    name = "character", 
    age = "numeric",
    email = "character",
    role = "character"
  ), 
  prototype = list(
    name = NA_character_,
    age = NA_real_,
    email = NA_character_,
    role = NA_character_
  )
)

setValidity("Person2", function(object) {
  if (length(object@name) != 1 | length(object@age) != 1 | length(object@email != 1)) {
    "@name, @age, and @email must be the same length"
  } else if (sum(object@roles %in% c("aut", "com", "cph", "cre", "ctb", "ctr", "dtc", "fnd", "rev", "ths", "trl")) != length(object@roles)) {
    "must be valid roles in @roles"
  } else {
    TRUE
  }
})
```

2. What happens if you define a new S4 class that doesn't have any slots? (Hint:
read about virtual classes in `?setClass`.)

A virtual class is created. Not sure how they can be used or are useful.

3. Imagine you were going to reimplement factors, dates, and data frames in S4. Sketch
out the `setClass()` calls that you would use to define the classes. Think about appropriate
`slots` and `prototype`.

```{r}
setClass("Factor",
  slots = c(
    data = "integer",
    levels = "character"),
  prototype = list(
    data = NA_integer_,
    levels = NA_character_)
)
```

```{r}
setClass("Dates",
  slots = c(
    date = "integer"
    ),
  prototype = list(
    date = NA_integer_
  )
)
```

```{r}
setClass("DataFrame",
  slots = c(
    data = "list",
    row.names = "character"
    ),
  prototype = list(
    data = list(),
    row.names = NA_character_
  )
)

```

# 15.4 Generics and methods

```{r}
# create S4 generic with setGeneric() that calls standardGeneric()
setGeneric("myGeneric", function(x) standardGeneric("myGeneric"))
```

* bad practice to use `{}` with `setGeneric()`

## 15.4.1 Signature

* `signature` allows user to control arguments that are used for method dispatch;
if not supplied, all args are used

```{r}
setGeneric("myGeneric", 
  function(x, ..., verbose = TRUE) standardGeneric("myGeneric"),
  signature = "x" # only use x
)
```

## 15.4.2 Methods

* S4 methods defined with `setMethod()`: need to define name of generic, name of
class, and method

```{r, eval = FALSE}
setMethod("myGeneric", "Person", function(x) {
  # method implementation
})
```

* second argument in `setMethod()` is the signature

* `methods("generic")` or `methods(class = "class")` to list methods that
belong to a generic or associated with a class

* `selectMethod("generic", "class")` to find the implementation of a specific
method

## 15.4.3 Show method

* `show()` controls how an object appears when printed

```{r}
args(getGeneric("show"))
```

```{r}
setMethod("show", "Person", function(object) {
  cat(is(object)[[1]], "\n",
      "  Name: ", object@name, "\n",
      "  Age:  ", object@age, "\n",
      sep = ""
  )
})
john
```

## 15.4.4 Accessors

* accessors return user-accessible slots

```{r}
person_name <- function(x) x@name
```

```{r}
setGeneric("name", function(x) standardGeneric("name"))
setMethod("name", "Person", function(x) x@name)

name(john)
```

* also provide a setter function if the slot is writable

```{r}
setGeneric("name<-", function(x, value) standardGeneric("name<-"))
setMethod("name<-", "Person", function(x, value) {
  x@name <- value
  validObject(x) # check if x is a valid object
  x
})

name(john) <- "Jon Smythe"
name(john)

#name(john) <- letters
```

## 15.4.5 Exercises

1. Add `age()` accessors for the `Person` class.

```{r}
setGeneric("age", function(x) standardGeneric("age"))
setMethod("age", "Person", function(x) x@age)

age(john)
```


2. In the definition of the generic, why is it necessary to repeat the name
of the generic twice?

`setGeneric()` creates the generic of the function while `standardGeneric`
initiates method dispatch.

3. Why does the `show()` method defined in Section 15.4.3 use `is(object)[[1]]`?
(Hint: try printing the employee subclass.)

```{r}
is(new("Employee"))

# if a subclass is provided, the class it inherits from is also printed
# take is(new("Employee"))[[1]] to ensure the subclass is only checked for
```

4. What happens if you define a method with different argument names to the generic?

```{r}
setClass("Person3", 
  slots = c(
    name = "character", 
    age = "numeric"
  ), 
  prototype = list(
    name = NA_character_,
    age = NA_real_
  )
)
```

```{r}
setGeneric("introduce", 
  function(x) standardGeneric("introduce")
)

setMethod("introduce", "Person3", function(x) {
  cat("My name is", x@name, "and I am", x@age, "years old.")
})

hadley <- new("Person3", name = "Hadley", age = 999)
introduce(hadley)
```

```{r}
setGeneric("introduce2", 
  function(x) standardGeneric("introduce2")
)

setMethod("introduce2", "Person3", function(obj) {
  cat("My name is", obj@name, "and I am", obj@age, "years old.")
})

hadley2 <- new("Person3", name = "Hadley", age = 999)
introduce2(hadley)
# a warning is produced
```

# 15.5 Method dispatch

* S4 important features: multiple inheritance and multiple dispatch

* recommended to keep dispatch simple by avoiding multiple inheritance and
using multiple dispatch when absolutely necessary

## 15.5.1 Single dispatch

* `f(...)` defines scope (ie: `f(;P -> ;0 -> :0)`)

* method graph displays all possible methods that can be defined
(ie: ;P -> ;0 -> | :0 | ), where emoji in | has been defined with
`setMethod()`

* least general class -> most general class; if no method is found, method
dispatch fails and error is thrown

* always define methods for terminal nodes! (most general classes)

* pseudo-classes don't exist but allow to define useful behaviors

- `ANY` matches to any class (link to `ANY` is longer than links between
other classes)

- ie: ;P -> ;0 -> :0 ---> ANY

- `MISSING` matches whenever the argument is missing; important for
`+` and `-` that use double dispatch

## 15.5.2 Multiple inheritance

* `f(B) -> B)` & `f(B) -> :) -> :0)`

* closest method is selected when finding multiple methods

* actual details in multiple inheritance in S4 in `?Methods_Details`

* when methods are the same distance, it is an ambiguous method
(drawn with || B) ||); ambiguity is resolved by providing more
precise methods

* `ANY` will never contribute to ambiguity since it is further away

## 15.5.5 Exercises

1. lol no

2. no thanks

3. nope

# 15.6 S4 and S3

## 15.6.1 Classes

* `slots` and `contains` can use S4, S3, or implicit classes

* must register with `setOldClass()` if using S3 class

```{r}
setOldClass("data.frame")
setOldClass(c("ordered", "factor"))
setOldClass(c("glm", "lm"))
```

```{r}
# better to provide full S4 definition
setClass("factor",
  contains = "integer",
  slots = c(
    levels = "character"
  ),
  prototype = structure(
    integer(),
    levels = character()
  )
)
setOldClass("factor", S4Class = "factor")
```

* if S4 object inherits from S3 class, has `.Data` virtual slot

```{r}
RangedNumeric <- setClass(
  "RangedNumeric",
  contains = "numeric",
  slots = c(min = "numeric", max = "numeric"),
  prototype = structure(numeric(), min = NA_real_, max = NA_real_)
)
rn <- RangedNumeric(1:10, min = 1, max = 10)
rn@min
rn@.Data
```

## 15.6.2 Generics

* can convert S3 generics to S4 generic

```{r}
setGeneric("mean")
selectMethod("mean", "ANY") # existing method becomes default method
```

## 15.6.3 Exercises

1. What would a full `setOldClass()` definition look like for an
ordered factor (ie: add `slots` and `prototype` the definition above)?

```{r}
Factor <- setClass(
  "Factor",
  contains = "factor",
  slots = c(levels = "character"),
  prototype = structure(integer(), levels = character())
)
```

2. Define a `length` method for the `Person` class.

```{r, eval = FALSE}
setClass("Person4", 
  slots = c(
    name = "character", 
    age = "numeric",
    length = "numeric"
  ), 
  prototype = list(
    name = NA_character_,
    age = NA_real_,
    length = NA_real_
  )
)

setGeneric("length", function(x) standardGeneric("length"))
setMethod("length", "Person4", function(x) x@length)

setGeneric("length<-", function(x, value) standardGeneric("length<-"))
setMethod("length<-", "Person4", function(x, value) {
  x@length <- value
  validObject(x) # check if x is a valid object
  x
})
```















