---
title: 'Chapter 11: Function operators'
author: "Paulo Magalang"
date: "2023-02-21"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(purrr)
library(memoise)
```

# 11.1 Introduction

* function operator: function that takes one (or more) functions as input and returns
a function as an output

```{r}
chatty <- function(f) {
  force(f)
  
  function(x, ...) {
    res <- f(x, ...)
    cat("Processing ", x, "\n", sep = "")
    res
  }
}
f <- function(x) x ^ 2
s <- c(3, 2, 1)

purrr::map_dbl(s, chatty(f))
```

# 11.2 Existing function operators

## 11.2.1 Capturing errors with `purrr::safely()`

```{r}
x <- list(
  c(0.512, 0.165, 0.717),
  c(0.064, 0.781, 0.427),
  c(0.890, 0.785, 0.495),
  "oops"
)

#out <- rep(NA_real_, length(x))
#for (i in seq_along(x)) {
#  out[[i]] <- sum(x[[i]])
#}
#out
```

* `purrr::safely()`: function operator that transforms a functio to turn errors into data

```{r}
safe_sum <- safely(sum)
safe_sum
```

```{r}
str(safe_sum(x[[1]]))
str(safe_sum(x[[4]]))
```

```{r}
#out <- map(x, safely(sum))
#str(out)
```

```{r}
out <- transpose(map(x, safely(sum)))
str(out)
```

```{r}
ok <- map_lgl(out$error, is.null)
ok

x[!ok]
out$result[ok]
```

```{r, eval = FALSE}
fit_model <- function(df) {
  glm(y ~ x1 + x2 * x3, data = df)
}

models <- transpose(map(datasets, safely(fit_model)))
ok <- map_lgl(models$error, is.null)

# which data failed to converge?
datasets[!ok]

# which models were successful?
models[ok]
```

## 11.2.2 Caching computations with `memoise::memoise()`

* memoization of a function: a function remembers previous inputs and returns cached results

```{r}
slow_function <- function(x) {
  Sys.sleep(1)
  x * 10 * runif(1)
}
system.time(print(slow_function(1)))
system.time(print(slow_function(1)))
```

```{r}
fast_function <- memoise::memoise(slow_function)

system.time(print(fast_function(1)))
system.time(print(fast_function(1)))
```

```{r}
fib <- function(n) {
  if (n < 2) return(1)
  fib(n - 2) + fib(n - 1)
}
system.time(fib(23))
system.time(fib(24))
```

```{r}
fib2 <- memoise::memoise(function(n) {
  if (n < 2) return(1)
  fib2(n - 2) + fib2(n - 1)
})
system.time(fib2(23))
system.time(fib2(24))
```

## 11.2.3 Exercises

1. Base R provides a function operator in the form of `Vectorize()`. What does it do?
When might you use it?

`Vectorize()` allows you to input vectors to functions that expect single values.

```{r}
# example from documentation
vrep <- Vectorize(rep.int)
vrep(1:4, 4:1)

# equivalent to rep.int(x[i], times = y[j]) and stores as list
```


2. Read the source code for `possibly()`. How does it work?

```
function (.f, otherwise = NULL, quiet = TRUE) 
{
    .f <- as_mapper(.f)
    force(otherwise)
    check_bool(quiet)
    function(...) {
        tryCatch(.f(...), error = function(e) {
            if (!quiet) 
                message("Error: ", conditionMessage(e)) # outputs a message instead of throwing an error
            otherwise
        })
    }
}
```

3. Read the source code for `safely()`. How does it work?

```
function (.f, otherwise = NULL, quiet = TRUE) 
{
    .f <- as_mapper(.f)
    force(otherwise)
    check_bool(quiet)
    function(...) capture_error(.f(...), otherwise, quiet)
}
```

# 11.3 Case study: Creating your own function operators

```{r}
urls <- c(
  "adv-r" = "https://adv-r.hadley.nz", 
  "r4ds" = "http://r4ds.had.co.nz/"
  # and many many more
)
path <- paste(tempdir(), names(urls), ".html")

walk2(urls, path, download.file, quiet = TRUE)
```

```{r}
# add delay
delay_by <- function(f, amount) {
  force(f)
  force(amount)
  
  function(...) {
    Sys.sleep(amount)
    f(...)
  }
}
system.time(runif(100))
system.time(delay_by(runif, 0.1)(100))

walk2(urls, path, delay_by(download.file, 0.1), quiet = TRUE)
```

```{r}
dot_every <- function(f, n) {
  force(f)
  force(n)
  
  i <- 0
  function(...) {
    i <<- i + 1 # add internal counter
    if (i %% n == 0) cat(".")
    f(...)
  }
}
walk(1:100, runif)
walk(1:100, dot_every(runif, 10))
```

```{r}
walk2(
  urls, path, 
  dot_every(delay_by(download.file, 0.1), 10), 
  quiet = TRUE
)
```

```{r}
walk2(
  urls, path, 
  download.file %>% dot_every(10) %>% delay_by(0.1), 
  quiet = TRUE
)
```

## 11.3.1 Exercises

1. Weigh the pros and cons of `download.file %>% dot_every(10) %>% delay_by(0.1)` vs
`download.file %>% delay_by(0.1) %>% dot_every(10)`.

If the delay is large, the first pipeline is preferred since the user will have to wait
more time to see if the code is actually running.

2. Should you memoize `download.file()`? Why or why not?

Depends on the size of the file being downloaded. You would need at most two times the
memory to download a file if downloading files is memoized. Even if the files are small,
memoizing file downloads is unnecessary since those files will probably download faster
anyways. It might take longer since you would have to open and write many small files.

3. Create a function operator that reports whenever a file is created or deleted in the
working directory, using `dir()` and `setdiff()`. What other global function effects might you
want to track?

```{r}
# inputs are vectors
compare_dir <- function(old_dir, new_dir) {
  force(old_dir)
  force(new_dir)
  
  # check whether files were created or destroyed by comparing vector lengths
  if (length(old_dir) > length(new_dir)) { # file destroyed
    cat(paste0(setdiff(old_dir, new_dir), " deleted\n"))
  } else if (length(old_dir) < length(new_dir)) { # file created
    cat(paste0(setdiff(new_dir, old_dir), " created\n"))
  } else { # no change
    cat("No changes")
  }
}

# create function operator
check_dir <- function(f) {
  force(f)
  
  function(...) { # does this function need any input? no
    # need to get current directory files
    current_dir <- dir()
    
    # call the function f() and compare_dir() at some point
    f(...) # execute function
    changed_dir <- dir()
    
    compare_dir(current_dir, changed_dir)
  }
}
```

```{r}
# took test cases from Advanced R solutions
file_create <- check_dir(file.create)
file_remove <- check_dir(file.remove)

file_create("delete_me")
file_remove("delete_me")
```

4. Write a function operator that logs a timestamp and message to a file every time a function
is run.

```{r}
get_timestamp <- function(...) {
  time_date <- Sys.time() %>% as.character()
  return(time_date)
}

print_to_file <- function(file_to_write, msg) {
  if (!file.exists(file_to_write)) {
    file.create(file_to_write)
  }
  
  write(msg, file = file_to_write, append = TRUE)
}

log_function <- function(f, file_to_write = "log.txt") {
  force(f)
  force(file_to_write)
  
  function(...) {
    time <- get_timestamp()
    msg <- paste0(time, ": function called")
    
    print_to_file(file_to_write, msg)
    f(...)
  }
}
```

```{r}
# took test cases from Advanced R solutions
mean2 <- log_function(mean)
mean2(1:4)
Sys.sleep(1)
mean2(1:4)
```

5. Modify `delay_by()` so that instead of delaying by a fixed amount of time, it ensures that
a certain amount of time has elapsed since the function was last called. That is, if you called
`g <- delay_by(1, f); g(); Sys.sleep(2); g()` there shouldn't be an extra delay.

```{r}
delay_by <- function(f, amount) {
  force(f)
  force(amount)
  
  function(...) {
    Sys.sleep(amount)
    f(...)
  }
}
```

```{r}
delay_by2 <- function(f, amount) {
  force(f)
  force(amount)
  
  time <- NULL
  
  function(...) {
    if (!is.null(time)) { # if not init step, just run function
      execution_time <- Sys.time()
      
      time_diff <- abs(time - execution_time)
      if (as.numeric(time_diff) < amount) { # if time between executions below min threshold, delay
        wait <- abs(time_diff - amount)
        Sys.sleep(wait)
        #Sys.sleep(amount)
      }
    }
    
    time <<- Sys.time()
    f(...)
  }
}
```

```{r}
runif2 <- delay_by2(runif, 4)

system.time(runif2(100))
Sys.sleep(2)
system.time(runif2(100))
```













